---
title: "ybcu_kern_plant_plots"
author: "Patrick D. lorch"
date: "2025-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bring in spreadsheet data


### 2024

This assumes output from ArcGIS fieldmaps.

Tallys live and dead hits at each meter of height at each meter along a cardinal direction transect.

```{r dataingest24}
library(dplyr)
# install.packages(rio)
library(rio)
library(qdapTools)

data_list_2024 = import_list(file.path("C:",
                        "Users",
                        "PatrickLorch",
                        "SSRS",
                        "Southern Sierra Research Station - Documents",
                        "Projects",
                        "YBCU - Other_Kern",
                        "Data",
                        "2024",
                        "YBCU Nest Veg 2024-needs foliage density compilation.xlsx")) # , setclass = "tbl", , which = c(1, 2)

names(data_list_2024)
View(data_list_2024$E_Foliage_Density)

e_density_24 = data_list_2024$E_Foliage_Density
w_density_24 = data_list_2024$W_Foliage_Density
n_density_24 = data_list_2024$N_Foliage_Density
s_density_24 = data_list_2024$S_Foliage_Density
# names(e_density_24)

e_density_24_hits = e_density_24 %>%
  select(1:109)
w_density_24_hits = w_density_24 %>%
  select(1:109)
n_density_24_hits = n_density_24 %>%
  select(1:109)
s_density_24_hits = s_density_24 %>%
  select(1:109)

base_names = names(e_density_24_hits)[c(1:4, 105:109)]
hit_col_names = names(e_density_24_hits)[5:104]
heights = unique(sapply(strsplit(hit_col_names," "), `[`, 2))

e_density_24_hits_sum = e_density_24_hits %>%
  select(contains(base_names))
```

## Tally hits 

Tally number of 1m heights that have veg at each 1m stop along transects in each cardinal direction.

```{r hittally}

hit_tally = function(df = e_density_24_hits, hght){
  df %>%
    select(contains(hght)) %>%
    t() %>%
    as_tibble() %>%
    qdapTools::mtabulate() %>%
    rename_with(~ paste0(hght, " ", .x, recycle0 = T))
}

e_density_24_hits_sum = 
  e_density_24_hits %>%
     mutate(hit_tally(., hght = heights[1]),
            hit_tally(., hght = heights[2]),
            hit_tally(., hght = heights[3]),
            hit_tally(., hght = heights[4]),
            hit_tally(., hght = heights[5]),
            hit_tally(., hght = heights[6]),
            hit_tally(., hght = heights[7]),
            hit_tally(., hght = heights[8]),
            hit_tally(., hght = heights[9]),
            hit_tally(., hght = heights[10])) %>%
  select(contains(base_names), ends_with(c(" X", " D"))) %>%
  rename_with(~ gsub(" X", " live", .x, fixed = T)) %>%
  rename_with(~ gsub(" D", " dead", .x, fixed = T))

write.csv(e_density_24_hits_sum, "e_density_24_hits_sum.csv", row.names = F)

w_density_24_hits_sum = 
  w_density_24_hits %>%
     mutate(hit_tally(., hght = heights[1]),
            hit_tally(., hght = heights[2]),
            hit_tally(., hght = heights[3]),
            hit_tally(., hght = heights[4]),
            hit_tally(., hght = heights[5]),
            hit_tally(., hght = heights[6]),
            hit_tally(., hght = heights[7]),
            hit_tally(., hght = heights[8]),
            hit_tally(., hght = heights[9]),
            hit_tally(., hght = heights[10])) %>%
  select(contains(base_names), ends_with(c(" X", " D"))) %>%
  rename_with(~ gsub(" X", " live", .x, fixed = T)) %>%
  rename_with(~ gsub(" D", " dead", .x, fixed = T))

write.csv(w_density_24_hits_sum, "w_density_24_hits_sum.csv", row.names = F)

n_density_24_hits_sum = 
  n_density_24_hits %>%
     mutate(hit_tally(., hght = heights[1]),
            hit_tally(., hght = heights[2]),
            hit_tally(., hght = heights[3]),
            hit_tally(., hght = heights[4]),
            hit_tally(., hght = heights[5]),
            hit_tally(., hght = heights[6]),
            hit_tally(., hght = heights[7]),
            hit_tally(., hght = heights[8]),
            hit_tally(., hght = heights[9]),
            hit_tally(., hght = heights[10])) %>%
  select(contains(base_names), ends_with(c(" X", " D"))) %>%
  rename_with(~ gsub(" X", " live", .x, fixed = T)) %>%
  rename_with(~ gsub(" D", " dead", .x, fixed = T))

write.csv(n_density_24_hits_sum, "n_density_24_hits_sum.csv", row.names = F)

s_density_24_hits_sum = 
  s_density_24_hits %>%
     mutate(hit_tally(., hght = heights[1]),
            hit_tally(., hght = heights[2]),
            hit_tally(., hght = heights[3]),
            hit_tally(., hght = heights[4]),
            hit_tally(., hght = heights[5]),
            hit_tally(., hght = heights[6]),
            hit_tally(., hght = heights[7]),
            hit_tally(., hght = heights[8]),
            hit_tally(., hght = heights[9]),
            hit_tally(., hght = heights[10])) %>%
  select(contains(base_names), ends_with(c(" X", " D"))) %>%
  rename_with(~ gsub(" X", " live", .x, fixed = T)) %>%
  rename_with(~ gsub(" D", " dead", .x, fixed = T))

write.csv(s_density_24_hits_sum, "s_density_24_hits_sum.csv", row.names = F)
  
  
```


## Join in plot data

This is an example of how you would join  plot data to other subtables for analyses that require this.

Just show an example of 

```{r joinplot}
library(dplyr)

veg_plots_24 = data_list_2024$Vegetation_Sampling_Plots
names(veg_plots_24)

veg_den_plot_join = function(veg = veg_plots_24,
                             den = e_density_24_hits_sum){
  veg %>%
  rename(OBJECTID_plot = OBJECTID,
         CreationDate_plot = CreationDate,
         LastEditDate_plot = LastEditDate,
         Longitude = x,
         Latitude = y) %>%
  left_join(den, by = join_by(GlobalID == "Parent GUID"))
}

e_density_24_hits_sum_w_plot = 
  veg_den_plot_join(veg = veg_plots_24,
                    den = e_density_24_hits_sum)

w_density_24_hits_sum_w_plot = 
  veg_den_plot_join(veg = veg_plots_24,
                    den = w_density_24_hits_sum)

n_density_24_hits_sum_w_plot = 
  veg_den_plot_join(veg = veg_plots_24,
                    den = n_density_24_hits_sum)

s_density_24_hits_sum_w_plot = 
  veg_den_plot_join(veg = veg_plots_24,
                    den = s_density_24_hits_sum)

```

